Measuring Hash Trustworthiness via Collision Utility Metrics: Logical Cryptanalysis of MD4

### 2017-11-28 dsn-2018/bare_conf.tex
Flatten subsubsections, expand techniques
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

% *** CITATION PACKAGES ***
%
\usepackage{cite}
\usepackage{hyperref,supertabular,amssymb}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.

@@

$\mathcal{C}$ for collision class. Ordered tuple.
$\mathcal{F}$ for family of collision classes. Set.
$I$ is an indexing function.

\section{The Utility of a Collision}

@@

We introduce a distance function, $\delta$ between collision classes by the
number of differences in intermediate rounds deltas. That is, given two
collision classes, $C_1, C_2 \in \mathcal{C}$:
\begin{align}
    \delta(C_1, C_2) = \big| \{ i : I(C_1, i) \neq I(C_2, i) \} \big|
\end{align}
We find justification for this metric in the existing literature on MD4:
H. Dobbertin's \cite{Dobbertin1998} and P. Kasselman's \cite{KasselmanMD4}
collision classes have distance $1$ under this metric.

Please refer to Table \ref{table:distance} for the distances between existing
collisions in MD4.

@@

We define the neighborhood of a collision class, $C \in \mathcal{C}$, to be the
set of all other collisions at a fixed distance, $d \in \mathbb{N}$, from $C$.
That is:
\begin{align}
    N(C, d) & = \{ C_j : \delta(C, C_j) = d \} \\
    N(C) & = N(C, 1)
\end{align}
For instance, $C_{Wang} \in N(C_{Sasaki}, 21)$. The distance parameter, $d$,
may optionally be ommited, in which case the unit distance is implied. Thus,
$C_{Kasselman} \in N(C_{Dobbertin})$.

The implications of neighborhoods are discussed further in section
\ref{empirical:neighborhood}.

We can similarly define the neighborhood of a family of collision classes,
$F \in \mathcal{F}$, to be the set of all other families of collision classes
at a fixed distance, $d \in \mathbb{N}$, from $F$.

\subsection{Reduced Rounds}
\subsubsection{Abstract}

\textbf{TODO} - Describe technique applied to tags.

\subsubsection{Neighborhood}

@@

@@

\textbf{TODO}
Please refer to Table \ref{table:differentials} for the number of differential
paths...

@@

Please refer to Table \ref{table:neighborhood} for the sizes of neighborhoods
of existing collisions. In particular, note that Y. Sasaki's attack--which
claimed to improve upon X. Wang's attack--has a larger neighborhood size, and
likewise with P. Kasselman's attack which improved upon H. Dobbertin's attack.

@@

>>>>

% *** CITATION PACKAGES ***
%
\usepackage{cite}
\usepackage[hidelinks]{hyperref}
\usepackage{supertabular,amssymb}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.

@@

$\mathcal{C}$ for collision class. Ordered tuple, indexable.
$\mathcal{F}$ for family of collision classes. Ordered set, indexable.
$[i]$ is an indexing function.
$R$ is number of rounds of $F$ or $C$.

\section{Measuring the Utility of a Collision}

@@

We introduce a distance function, $\delta$ between collision classes by the
number of differences in intermediate rounds deltas. That is, given two
collision classes, $C_1, C_2 \in \mathcal{C}$:
\begin{align}
    \delta(C_1, C_2) = \big| \{ i : C_1[i] \neq C_2[i] \} \big|
\end{align}
We find justification for this metric in the existing literature on MD4:
H. Dobbertin's \cite{Dobbertin1998} and P. Kasselman's \cite{KasselmanMD4}
collision classes have distance $1$ under this metric.

Refer to Table \ref{table:distance} for the distances between existing
collisions in MD4.

@@

We define the neighborhood of a collision class, $C \in \mathcal{C}$, to be the
set of all other collisions at a fixed distance, $d \in \mathbb{N}$, from $C$.
That is:
\begin{align}
    N(C, d) & = \{ C_j \in \mathcal{C} : \delta(C, C_j) = d \} \\
    N(C) & = N(C, 1)
\end{align}
For instance, $C_{Wang} \in N(C_{Sasaki}, 21)$. The distance parameter, $d$,
may optionally be omitted, in which case the unit distance is implied. Thus,
$C_{Kasselman} \in N(C_{Dobbertin})$.

The implications of collision class neighborhoods are discussed further in
section \ref{empirical:neighborhood}.

We can similarly define the neighborhood of a family of collision classes,
$F \in \mathcal{F}$, to be the set of all other families of collision classes
at a fixed distance, $d \in \mathbb{N}$, from $F$. That is:
\begin{align}
    \mathcal{N}(F, d) & = \{ F_j \in \mathcal{F} : \delta(F, F_j) = d \} \\
    \mathcal{N}(F) & = N(F, 1)
\end{align}
The distance parameter, $d$, may optionally be omitted, in which case the unit
distance is implied.

Neighborhoods can be classified into three types: \textit{expansion},
\textit{internal}, and \textit{mixed}. Let $d$ be fixed. An \textit{expansion}
neighborhood of a collision class, $C \in \mathcal{C}$, is the neighborhood
restricted only to those collision classes which only differ in rounds external
to the collision family of $C$. An \textit{internal} neighborhood of $C$ is the
neighborhood restricted only to those collision classes which differ in rounds
internal to the collision family of $C$. An \textit{mixed} neighborhood of $C$
is the neighborhood restricted only to collisions which differ in a round
internal and a round external to the collision family of $C$. That is:
\begin{align*}
    N_{exp}(C, d) = \{ & C_{j} \in N(C, d) : \forall i \in F(C), C_j[i] = C[i] \} \\
    N_{int}(C, d) = \{ & C_{j} \in N(C, d) : \forall i \not\in F(C), C_j[i] = C[i] \} \\
    N_{mix}(C, d) = \{ & C_{j} \in N(C, d) : \exists i \in F(C), C_j[i] \neq C[i] \\
        & \text{ and } \exists k \not\in F(C), C_j[k] \neq C[k] \}
\end{align*}

\subsection{Family Similarity}
We define a relation among families of collisions across rounds. Let $F_1$ and
$F_2$ be two different families of collisions. Then we say that $F_1$ and $F_2$
are \textit{similar}, and notate it $F_1 \lesssim F_2$, if $R(F_1) \leq R(F_2)$
and $F_1 \subseteq F_2$. Note that, when $R(F_1) = R(F_2)$, this is equivalent
to saying that $F_2$ is in some expansion neighborhood of $F_1$. Further, when
$R(F_1) < R(F_2)$ and $F_1 = F_2$, then we say that $F_2$ is the
\textit{trivial extension} of $F_1$.

We claim that the following statements are true:
\begin{enumerate}
    \item For every $F \in \mathcal{F}$, there exists $F'$ such that
        $F' \lesssim F$ and $R(F') + 4 = R(F)$.
    \item For every $F \in \mathcal{F}$, there exists $F'$ such that
        $F \lesssim F'$.
\end{enumerate}

\textbf{TODO} - Describe justification? Later sections?

\subsection{Class Similarity}

We claim that the aforementioned statements hold when considering individual
collision classes instead of families of collision classes, but with less
likelihood.

@@

\subsection{Miscellaneous}

\textbf{TODO} - Describe chosen prefix, second preimage, ASCII, and other
tips and tricks for faster model runs.

@@

\textbf{TODO}
Refer to Table \ref{table:differentials} for the number of differential
paths...

@@

Refer to Table \ref{table:neighborhood} for the sizes of neighborhoods
of existing collisions. In particular, note that Y. Sasaki's attack--which
claimed to improve upon X. Wang's attack--has a larger neighborhood size, and
likewise with P. Kasselman's attack which improved upon H. Dobbertin's attack.

@@

\section{New Collision}

\section{Broader Impacts}

\textbf{TODO} - Discuss impacts on real-world systems like MD5, SHA-1, etc.
